<<<<<<< HEAD
--
title : "Personal Activity Monitoring Anaysis"
Author: "Nathan Russell"
--
## Mean total steps taken per day
```{r load-data, echo=TRUE}
library(dplyr)
library(ggplot2)
#Load data
activity_data <- read.csv("activity.csv")
activity_data$date <- as.Date(activity_data$date, format="%Y-%m-%d")
str(activity_data)
# Sum total steps per day
total_steps_per_day <- activity_data%>%
  group_by(date) %>%
  summarize(total_steps = sum(steps, na.rm = TRUE))
# Display first rows
head(total_steps_per_day)
# plot histogram
ggplot(total_steps_per_day, aes(x = total_steps)) + 
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(title = "Histogram of Total steps per day",
       x = "Total Steps",
       y = "Frequency")
#Calculate Mean and median
mean_steps <- mean(total_steps_per_day$total_steps, na.rm = TRUE)
median_steps <- median(total_steps_per_day$total_steps, na.rm = TRUE)
# display results 
mean_steps
median_steps
```

## Average Daily Activity Pattern
```{r avg-daily-pattern, echo=TRUE}
# find avg number of steps over 5 mintue interval
avg_steps_interval <- activity_data %>% 
  group_by(interval) %>%
  summarize(mean_steps = mean(steps, na.rm =TRUE)) #compute mean steps
  
head(avg_steps_interval)

### Time Series plot 
ggplot(avg_steps_interval, aes(x = interval, y = mean_steps)) + 
  geom_line(color = "blue", linewidth = 1) +
  labs(title = "Average Daily Activity Pattern",
       x = "5-mintue Interval",
       y = "Average Steps")
## Max Interval 

max_interval <- avg_steps_interval %>%
  filter(mean_steps == max(mean_steps))
#display max interval
max_interval
```

## Handling missing values

```{r count-missing, echo=TRUE}
missing_values <- sum(is.na(activity_data$steps))
#display number of missing steps 
missing_values

# compute mean steps for each 5 min interval (ignore NA)
interval_means <- activity_data %>%
  group_by(interval) %>%
  summarize(mean_steps = mean(steps, na.rm = TRUE))

# Display the first few rows of the interval means
head(interval_means)

#merge interval means w dataset
activity_imputed <- activity_data %>%
  left_join(interval_means, by = "interval")
#replace NA values in steps columnn with mean for interval
activity_imputed <- activity_imputed %>%
  mutate(steps = ifelse(is.na(steps), mean_steps, steps)) %>%
  select(-mean_steps)
#count missing values after imputation
sum(is.na(activity_imputed$steps))

head(activity_imputed)
```

### Activity pattern differences between weekdays and weekends
```{r activity-differences, echo = TRUE}
activity_imputed$date <- as.Date(activity_imputed$date)
# create factor variable for weekday vs weekend
activity_imputed <- activity_imputed %>% 
  mutate(day_type = ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
#convert to factor variable
activity_imputed$day_type <- factor(activity_imputed$day_type, levels = c("weekday", "weekend"))
head(activity_imputed)
#head confirms appropriate classification
#compute avg number of steps for interval, separated by weekday/weekend
avg_steps_weekday_weekend <- activity_imputed %>%
  group_by(interval, day_type) %>%
  summarize(mean_steps = mean(steps, na.rm = TRUE))
#display first rows
head(avg_steps_weekday_weekend)

#create panel plot
ggplot(avg_steps_weekday_weekend, aes(x = interval, y = mean_steps)) +
  geom_line(color = "blue") +
  facet_wrap(~day_type, nrow = 2) +
  labs(title = "Average Daily Activity Pattern: Weekdays vs. Weekends",
       x = "5-Minute Interval",
       y = "Average Number of Steps")
```


=======
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
